\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{array}
\usepackage{longtable}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{listings}
\lstset{basicstyle=\ttfamily\small,breaklines=true,columns=fullflexible,frame=single}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.65em}
\title{Lab 2 Extension: Pipelining and Superscalar Design Space}
\author{Computer Architecture Laboratory}
\date{}
\begin{document}
\maketitle
\section{Introduction}
This lab extends the previous processor-design lab. Students change one architectural feature at a time and identify whether performance is limited by issue policy, pipeline width, functional units, or instruction window size.
Run setup commands from the cloned repository root:
\begin{lstlisting}
cd /path/to/SimpleScalar
make clean
make config-pisa
make
mkdir -p labs
cp tests-pisa/bin.little/test-math labs/
\end{lstlisting}
\section{Part 1: Increasing Functional Units in an In-Order Processor}
Create \texttt{config\_c.cfg}:
\begin{lstlisting}
cp config/default.cfg labs/config_c.cfg
\end{lstlisting}
In \texttt{labs/config\_c.cfg}, set:
\begin{lstlisting}
-decode:width 1
-fetch:ifqsize 1
-issue:width 1
-issue:inorder true
-ruu:size 8
-lsq:size 8
-res:ialu 4
-res:imult 4
-res:memport 4
-res:fpalu 4
-res:fpmult 4
\end{lstlisting}
Run:
\begin{lstlisting}
cd labs
../sim-outorder -config config_c.cfg -redir:sim sim_configc.out ./test-math
\end{lstlisting}
Questions:
\begin{enumerate}
\item What is the IPC of \texttt{test-math} using this configuration?
\item Compare this configuration with configuration A from Lab 2. Does adding resources increase performance in an in-order pipeline?
\end{enumerate}
\section{Part 2: Increasing Functional Units With Out-of-Order Issue}
Create \texttt{config\_d.cfg}:
\begin{lstlisting}
cp config_c.cfg config_d.cfg
\end{lstlisting}
Change:
\begin{lstlisting}
-issue:inorder false
\end{lstlisting}
Run:
\begin{lstlisting}
../sim-outorder -config config_d.cfg -redir:sim sim_configd.out ./test-math
\end{lstlisting}
Questions:
\begin{enumerate}
\item What is the IPC using this configuration?
\item Compare this configuration with Part 1. Does out-of-order execution help when resources are plentiful but pipeline width is narrow?
\item The system contains four functional units. Find the bottleneck by running multiple simulations based on \texttt{config\_d.cfg}, changing only one of these parameters to 4 at a time:
\end{enumerate}
\begin{lstlisting}
-decode:width 4
-issue:width 4
\end{lstlisting}
\begin{enumerate}
\item Does performance increase? What is the main bottleneck in configuration D?
\end{enumerate}
\section{Part 3: Widening the Pipeline}
Create \texttt{config\_e.cfg}:
\begin{lstlisting}
cd /path/to/SimpleScalar
cp config/default.cfg labs/config_e.cfg
cd labs
\end{lstlisting}
In \texttt{config\_e.cfg}, set:
\begin{lstlisting}
-fetch:ifqsize 4
-decode:width 4
-lsq:size 8
-ruu:size 8
-issue:width 4
-issue:inorder false
-res:ialu 1
-res:imult 1
-res:memport 1
-res:fpalu 1
-res:fpmult 1
\end{lstlisting}
Run:
\begin{lstlisting}
../sim-outorder -config config_e.cfg -redir:sim sim_confige.out ./test-math
\end{lstlisting}
Questions:
\begin{enumerate}
\item What is the IPC for this configuration?
\item Compare this configuration with configuration B from Lab 2. What is the effect of widening the pipeline front end?
\end{enumerate}
\section{Part 4: Wide Pipeline and Many Resources}
Create \texttt{config\_f.cfg}:
\begin{lstlisting}
cp config_e.cfg config_f.cfg
\end{lstlisting}
Set all resource types to 4:
\begin{lstlisting}
-res:ialu 4
-res:imult 4
-res:memport 4
-res:fpalu 4
-res:fpmult 4
\end{lstlisting}
Run:
\begin{lstlisting}
../sim-outorder -config config_f.cfg -redir:sim sim_configf.out ./test-math
\end{lstlisting}
Question:
\begin{enumerate}
\item What is the IPC for this configuration?
\end{enumerate}
\section{Part 5: Increasing the Instruction Window}
Create \texttt{config\_g.cfg}:
\begin{lstlisting}
cp config_f.cfg config_g.cfg
\end{lstlisting}
Change:
\begin{lstlisting}
-ruu:size 16
\end{lstlisting}
Run:
\begin{lstlisting}
../sim-outorder -config config_g.cfg -redir:sim sim_configg.out ./test-math
\end{lstlisting}
Question:
\begin{enumerate}
\item What is the IPC for this configuration?
\end{enumerate}
\section{Summary Table}
\begin{center}
\scriptsize
\begin{longtable}{|p{0.10\linewidth}|p{0.10\linewidth}|p{0.10\linewidth}|p{0.10\linewidth}|p{0.10\linewidth}|p{0.10\linewidth}|p{0.10\linewidth}|p{0.10\linewidth}|p{0.10\linewidth}|}
\hline
Config & Issue mode & Fetch queue & Decode width & Issue width & RUU & LSQ & Functional units & IPC \\
\hline
\endfirsthead
\hline
Config & Issue mode & Fetch queue & Decode width & Issue width & RUU & LSQ & Functional units & IPC \\
\hline
\endhead
A & in-order & 1 & 1 & 1 & 8 & 8 & one each &  \\
\hline
B & out-of-order & 1 & 1 & 1 & 8 & 8 & one each &  \\
\hline
C & in-order & 1 & 1 & 1 & 8 & 8 & four each &  \\
\hline
D & out-of-order & 1 & 1 & 1 & 8 & 8 & four each &  \\
\hline
E & out-of-order & 4 & 4 & 4 & 8 & 8 & one each &  \\
\hline
F & out-of-order & 4 & 4 & 4 & 8 & 8 & four each &  \\
\hline
G & out-of-order & 4 & 4 & 4 & 16 & 8 & four each &  \\
\hline
\end{longtable}
\normalsize
\end{center}
\end{document}
