\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{array}
\usepackage{longtable}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{listings}
\lstset{basicstyle=\ttfamily\small,breaklines=true,columns=fullflexible,frame=single}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.65em}
\title{SimpleScalar Computer Architecture Lab Experiments}
\author{Computer Architecture Laboratory}
\date{}
\begin{document}
\maketitle
This folder contains the student experiment handouts for the SimpleScalar computer architecture lab sequence. Each experiment is provided as a PDF for printing, a Markdown file for copying commands, and a LaTeX source file for editing.
Repository layout after cloning:
\begin{lstlisting}
SimpleScalar/
  benchmarks/
  config/
  labs/
  scripts/
  SimpleScalar/
    experiments/
  tests-alpha/
  tests-pisa/
  sim-fast.c
  sim-safe.c
  sim-profile.c
  sim-bpred.c
  sim-outorder.c
  sim-singlecycle.c
  Makefile
\end{lstlisting}
There is no nested \texttt{simplesim-3.0/} directory in this repository. Unless a lab explicitly says to enter \texttt{labs/}, run commands from the cloned repository root:
\begin{lstlisting}
cd /path/to/SimpleScalar
\end{lstlisting}
\section{Suggested Lab Sequence}
\begin{itemize}
\item Lab 0: setup, build targets, and simulator overview. Files: \path{00_setup_and_orientation.pdf}, \path{00_setup_and_orientation.md}. Simulator: all.
\item Lab 1: instruction set profiling and Alpha vs PISA comparison. Files: \path{01_isa_profiling_lab.pdf}, \path{01_isa_profiling_lab.md}. Simulator: \texttt{sim-profile}.
\item Lab 2: in-order and out-of-order pipelined processors. Files: \path{02_processor_design_lab.pdf}, \path{02_processor_design_lab.md}. Simulator: \texttt{sim-outorder}.
\item Lab 2 extension: superscalar bottleneck and design-space analysis.\\
Files: \path{03_processor_design_home_assignment.pdf}\\
\path{03_processor_design_home_assignment.md}\\
Simulator: \texttt{sim-outorder}.
\item Lab 3: branch predictor accuracy and CPI impact. Files: \path{04_branch_prediction_lab.pdf}, \path{04_branch_prediction_lab.md}. Simulators: \texttt{sim-bpred}, \texttt{sim-outorder}.
\item Lab 4: educational single-cycle processor model. Files: \path{05_singlecycle_processor_lab.pdf}, \path{05_singlecycle_processor_lab.md}. Simulator: \texttt{sim-singlecycle}.
\item Lab 5: compile custom C programs for PISA. Files: \path{06_compiling_custom_c_for_pisa.pdf}, \path{06_compiling_custom_c_for_pisa.md}. Simulators: \texttt{sim-safe}, \texttt{sim-profile}.
\end{itemize}
\section{Notes}
\begin{itemize}
\item Lab 1 uses Alpha and PISA builds. Rebuild with \texttt{make clean} before switching ISA targets.
\item Labs 2, 4, and 5 use PISA.
\item Lab 3 uses Alpha benchmarks for branch prediction, matching the original UMass SimpleScalar reference lab.
\item The local PISA cross compiler is installed under \path{toolchain/} by \path{scripts/install_pisa_cross_compiler.sh}; it is not committed to the repository.
\end{itemize}
\end{document}
